Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2009-12-03 ltx/BrainIDE.tex
adjusted paper to include state-space model derived in Sheffield

% can use linebreaks \\ within to get better formatting as desired
\title{Estimation of Spatiotemporal Integro-Difference Equations for Dynamical Neural Field Models}

@@

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behavior. It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology.
The model relates the average number of action potentials $f$ arriving at $r$ to the local post synaptic membrane voltage. The post-synaptic potentials generated at a neuronal population at location $r$ by action potentials arriving from all other connected populations at locations $r'$ can be described by
\begin{equation}\label{SpikesToPotential}
v\left( {r,t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)f\left( {r,t'} \right)dt'} + e\left( {r,t} \right).
\end{equation}
The term $e(r,t)$ can be considered the potential generated by unmodeled inputs and spontaneous background activity. The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
h(t) = \left\{ {\begin{array}{*{20}{c}}
   {\exp ( - \alpha t)} & {t \ge 0}  \\
   0 & {t < 0}  \\
\end{array}} \right.
\end{equation}
where $\alpha=\tau^{-1}$ and $\tau$ is the synaptic time constant. The synaptic response kernel can be more elaborate, using rise and fall time constants, however, this simple form serves our purposes. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
f\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)p\left( {v\left( {r',\bar t} \right)} \right)dr'}
\end{equation}
where $\bar t = t - d\left| {r - r'} \right|$, $d$ is the propagation delay of action potentials, $p$ is spiking rate of populations $r'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface.

@@

p\left( {{v_t}\left( {r'} \right)} \right) = \frac{{{\nu _0}}}{{1 + \exp \left( {\beta \left( {{v_0}\left( r \right) - {v_t}\left( {r'} \right)} \right)} \right)}}.

@@

v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)p\left( v\left( r',\bar t \right)\right)dr' dt'+e\left(r,t\right)

@@

L\left( {\frac{\partial }{{dt}}} \right)v\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)p\left( {v\left( {r',\bar t} \right)} \right)dr'}  + e\left( {r,t} \right),

@@

\frac{{dv\left( {r,t} \right)}}{{dt}} + \alpha{v\left( {r,t} \right)} = \int_\Omega  {w\left( {r,r'} \right)p\left( {v\left( {r',\bar t} \right)} \right)dr'}  +e\left( {r,t} \right).

@@

\subsection{Assumptions and Simplifications}
We assume that the propagation delay of action potentials is zero. This is a common assumption that is made when modeling intracortical circuits (\textcolor[rgb]{1.00,0.00,0.00}{refs}). As stated in Wilson and Cohen's original works, the time constant $\tau$ is much greater than the axonal propagation delays, therefore it is reasonable to approximate the the delays as being negligible. Recent studies have demonstrated the importance of having a fast, but finite, propagation velocity for models to make realistic and complete predictions of cortical dynamics~\cite{Atay2005}. The purpose of this paper, however, is not to precisely simulate complex cortical dynamics, but to create a useful platform where estimation of hidden states is possible for the purpose of detecting abnormal neural states with intentions of generating informed patient specific control sequences. Therefore, the model simply has to be `good enough' for this purpose. Starting with this simplified version of the general model is a logical choice. The next assumption we will make is spatial homogeneity such that synaptic interaction between neural populations is only dependant on their spatial separation, not on their spatial location. Again, this is a common assumption when dealing with intracortical connectivity (\textcolor[rgb]{1.00,0.00,0.00}{ref}). This assumption is not essential for our IDE estimation techniques, where a heterogeneous structure is possible. The IDE estimation is derived for a general heterogeneous structure however, for clarity of results we shall only present the heterogeneous case. We will assume the synaptic response kernel has the same shape (scaled by the connectivity kernel) irrespective of the respective presynaptic population. In this case, inhibition can be modeled by having a negative response kernel and excitation with a positive response kernel, where the connectivity kernel provides the (scaled) polarity. This reduces the model to a single field. We will also assume the activation function $p(.)$ is a Heaviside step function. This is in agreement with the work of Schiff and Sauer~\cite{schiff2008kalman}. With these assumptions, the model has the form
\begin{equation}\label{SimplifiedModel}
\frac{{dv\left( {r,t} \right)}}{{dt}} = \int_\Omega  {w\left( {r-r'} \right)p\left( {v\left( {r',t} \right)} \right)dr'}  - \alpha{v\left( {r,t} \right)} + e\left( {r,t} \right).
In order to make the problem trackable, in terms of model order, we assume that we can decompose the field (membranes voltages) into a set of $n_\phi$ Gaussian basis functions. We will also assume that we know the location and width of the basis function that is required to decompose the field. The state of the system is the height basis functions, rather than the membrane potential. Therefore, the number of basis functions required to decompose the field is equal to the number of states. Also, we assume that we can represent the connectivity by a regular grid of three connectivity Gaussian basis functions (corresponding to short, mid and long range connectivity) of known width. The height of the connectivity basis functions ($\theta$ parameter) is to be estimated.
\subsection{IDE Formulation}
For the state-space derivation, to create an estimator, the model can be reformulated like so
\begin{equation}\label{new_spatiotemporal_kernels}
\frac{{dv\left( {r,t} \right)}}{{dt}} = \int_\Omega  {k\left( {r - r'} \right)\tilde v\left( {r',t} \right) - \alpha \delta \left( {r - r'} \right)\bar v\left( {r,t} \right)dr'}  + e\left( {r,t} \right)
\begin{equation}\label{define_v_tilde}
\tilde v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
   {v\left( {r,t} \right)} & {v\left( {r,t} \right) \ge {v_0}}  \\
   0 & {otherwise}  \\
\end{array}} \right.
\begin{equation}\label{define_v_bar}
\bar v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
   {v\left( {r,t} \right)} & {v\left( {r,t} \right) < {v_0}}  \\
   0 & {otherwise}  \\
\end{array}} \right.
\begin{figure*}
\centering
\includegraphics[scale=0.3]{ExampleFieldDecomposition.eps}
\caption[ExampleFieldDecomposition]{Illustration of a one dimensional field decomposition using 5 Gaussian basis functions, each weighted an element of the state vector. Each basis function has a width of 0.01 and is uniformly positioned with a separation of 0.25 (with arbitrary units).}
\label{ExampleFieldDecomposition}
\end{figure*}

\section{State-Space Neural Field Model}
In this section we shall reconstruct the model in a state-space form that is a novel representation of the neural field equations. The benefit of restating the model in this format is that it becomes easier to simulate and allows the connectivity to be inferred from data. \begin{equation}
 v_{t+1}(\mathbf r)=\int_{\Omega} \alpha \delta(\mathbf r-\mathbf r')\bar v_t(\mathbf r)d \mathbf r'+\int_{\Omega} w(\mathbf r -\mathbf r')\tilde v_t(\mathbf r')d\mathbf r'+e_t(\mathbf r)
\label{eq:cortexfield}
decomposing the field and the kernel we have
\begin{equation}
\bar v_t(\mathbf r)=\boldsymbol \phi^{\top}(\mathbf r)\bar{\mathbf x}_t
\label{eq:cortexfieldlv0}
\begin{equation}
\tilde v_t(\mathbf r)=\boldsymbol \phi^{\top}(\mathbf r)\tilde{\mathbf x}_t=\nu_0
\label{eq:cortexfieldgv0}
\begin{equation}
 w(\mathbf r -\mathbf r')=\boldsymbol \theta^{\top}\psi(\mathbf r -\mathbf r')
\label{eq:kerneldecomposition}
Substituting (\ref{eq:cortexfieldlv0}), (\ref{eq:cortexfieldgv0}) and (\ref{eq:kerneldecomposition}) in (\ref{eq:cortexfield}) we get
\begin{equation}
 \boldsymbol \phi^{\top}(\mathbf r)\bar{\mathbf x}_{t+1}+\boldsymbol \phi^{\top}(\mathbf r)\tilde{\mathbf x}_{t+1}=
\int_\Omega\alpha\delta(\mathbf r- \mathbf r')\boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_td\mathbf r'+\int_\Omega\boldsymbol \theta^{\top}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')\tilde{\mathbf x}_td\mathbf r'+e_t(\mathbf r)
\label{eq:decomposedcortexfield}
\begin{equation}
 \boldsymbol \phi^{\top}(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\int_{\Omega}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')\tilde{\mathbf x}_td\mathbf r'+e_t(\mathbf r)
by defining
\begin{equation}
 \boldsymbol\Phi(\mathbf r)=\int_{\Omega}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')d\mathbf r'
\label{eq:convolution}
we have
\begin{equation}
 \boldsymbol \phi^{\top}(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)\tilde{\mathbf x}_t+e_t(\mathbf r)
\label{eq:cortexfieldwithPhi}
Premultiplying (\ref{eq:cortexfieldwithPhi}) by $\boldsymbol\phi(\mathbf r) $, and integrating over $\mathbf r$
\begin{equation}
\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \phi^{\top}(\mathbf r)d\mathbf s(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \phi^{\top}(\mathbf r)d\mathbf r \bar{\mathbf x}_{t}+\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)d\mathbf r\tilde{\mathbf x}_t+\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r
\label{eq:cortexfieldversion2}
\begin{equation}
 \boldsymbol\Psi_x(\mathbf r)=\int_{\Omega}\boldsymbol\phi(\mathbf r)\boldsymbol\phi^{\top}(\mathbf r)d\mathbf r
\label{eq:psix}
\begin{equation}
  \boldsymbol\Psi_{\theta}(\mathbf r)=\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)d\mathbf r
\label{eq:psitheta}
substituting (\ref{eq:psix}) and (\ref{eq:psitheta}) in (\ref{eq:cortexfieldversion2}) we get
\begin{equation}
\boldsymbol\Psi_x(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\boldsymbol\Psi_x(\mathbf r) \bar{\mathbf x}_{t}+\boldsymbol\Psi_{\theta}(\mathbf r)\tilde{\mathbf x}_t+\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r
pre-multiplying both sides by $\boldsymbol\Psi_x^{-1}$
\begin{equation}
\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1}=\alpha\bar{\mathbf x}_{t}+\boldsymbol\Psi_x^{-1}\boldsymbol\Psi_{\theta}\tilde{\mathbf x}_t+\boldsymbol\Psi_x^{-1}\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r
????????????????????????????????
\subsection{Observation equation}
 \mathbf y_t(\mathbf r)&=&\\
&=&\int_{\Omega}H(\mathbf r-\mathbf r')v_t(\mathbf r')d\mathbf r'+\boldsymbol\epsilon_t \\
&=&\int_{\Omega}H(\mathbf r-\mathbf r')\bar v_t(\mathbf r')d\mathbf r'+\int_{\Omega}H(\mathbf r-\mathbf r')\tilde v_t(\mathbf r')d\mathbf r'+\epsilon_t
\label{eq:observationequation}

 we decompose $ H(\mathbf r-\mathbf r')$ as
 \begin{equation}
  H(\mathbf r-\mathbf r')= \mathbf c^{\top}\mu(\mathbf r-\mathbf r')
\label{eq:decomposedobservationkernel}
\end{equation}
and substituting  (\ref{eq:cortexfieldlv0}), (\ref{eq:cortexfieldgv0}) and (\ref{eq:decomposedobservationkernel}) in (\ref{eq:observationequation}) we get
\begin{equation}
 \mathbf y_t(\mathbf r)=\int_{\Omega}\mathbf c^{\top}\mu(\mathbf r-\mathbf r')\phi^{\top}(\mathbf r')\bar{\mathbf x}_{t}d\mathbf r'+\int_{\Omega}\mathbf c^{\top}\mu(\mathbf r-\mathbf r')\phi^{\top}(\mathbf r')\tilde{\mathbf x}_{t}d\mathbf r'+\epsilon_t
\end{equation}
and finally we have
\begin{equation}
\mathbf y_t(\mathbf r)= \begin{bmatrix}\mathbf c^{\top}\mathbf M(\mathbf r)&\mathbf c^{\top}\mathbf M(\mathbf r)\end{bmatrix}\underline{\mathbf x}_t+\epsilon_t
\end{equation}
where
\begin{equation}
 \mathbf M(\mathbf r)=\int_{\Omega}\mu(\mathbf r-\mathbf r')\phi^{\top}(\mathbf r')d\mathbf r'
\end{equation}
Next we shall demonstrate the stability of the model and in the following section the estimator for the state vector and connectivity structure is derived.

>>>>

% can use linebreaks \\ within to get better formatting as desired
\title{Estimation of Intracortical Connectivity in a Dynamical Neural Field Model}

@@

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behavior. It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology.
The model relates the average number of action potentials $g$ arriving at $r$ to the local post synaptic membrane voltage. The post-synaptic potentials generated at a neuronal population at location $r$ by action potentials arriving from all other connected populations at locations $r'$ can be described by
\begin{equation}\label{SpikesToPotential}
v\left( {r,t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {r,t'} \right)dt'} + e\left( {r,t} \right).
\end{equation}
The term $e(r,t)$ can be considered the potential generated by unmodeled inputs and spontaneous background activity. The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
h(t) = \left\{ {\begin{array}{*{20}{c}}
   {\exp ( - \alpha t)} & {t \ge 0}  \\
   0 & {t < 0}  \\
\end{array}} \right.
\end{equation}
where $\alpha=\tau^{-1}$ and $\tau$ is the synaptic time constant. The synaptic response kernel can be more elaborate, using rise and fall time constants, however, this simple form serves our purposes. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
g\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}
\end{equation}
where $\bar t = t - d\left| {r - r'} \right|$, $d$ is the propagation delay of action potentials, $f$ is spiking rate of populations $r'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface.

@@

f\left( {{v_t}\left( {r'} \right)} \right) = \frac{{{\nu _0}}}{{1 + \exp \left( {\beta \left( {{v_0}\left( r \right) - {v_t}\left( {r'} \right)} \right)} \right)}}.

@@

v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' dt'+e\left(r,t\right)

@@

L\left( {\frac{\partial }{{dt}}} \right)v\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  + e\left( {r,t} \right),

@@

\frac{{dv\left( {r,t} \right)}}{{dt}} + \alpha{v\left( {r,t} \right)} = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  +e\left( {r,t} \right).

@@

%\subsection{Assumptions and Simplifications}
%We assume that the propagation delay of action potentials is zero. This is a common assumption that is made when modeling intracortical circuits (\textcolor[rgb]{1.00,0.00,0.00}{refs}). As stated in Wilson and Cohen's original works, the time constant $\tau$ is much greater than the axonal propagation delays, therefore it is reasonable to approximate the the delays as being negligible. Recent studies have demonstrated the importance of having a fast, but finite, propagation velocity for models to make realistic and complete predictions of cortical dynamics~\cite{Atay2005}. The purpose of this paper, however, is not to precisely simulate complex cortical dynamics, but to create a useful platform where estimation of hidden states is possible for the purpose of detecting abnormal neural states with intentions of generating informed patient specific control sequences. Therefore, the model simply has to be `good enough' for this purpose. Starting with this simplified version of the general model is a logical choice. The next assumption we will make is spatial homogeneity such that synaptic interaction between neural populations is only dependant on their spatial separation, not on their spatial location. Again, this is a common assumption when dealing with intracortical connectivity (\textcolor[rgb]{1.00,0.00,0.00}{ref}). This assumption is not essential for our IDE estimation techniques, where a heterogeneous structure is possible. The IDE estimation is derived for a general heterogeneous structure however, for clarity of results we shall only present the heterogeneous case. We will assume the synaptic response kernel has the same shape (scaled by the connectivity kernel) irrespective of the respective presynaptic population. In this case, inhibition can be modeled by having a negative response kernel and excitation with a positive response kernel, where the connectivity kernel provides the (scaled) polarity. This reduces the model to a single field. We will also assume the activation function $p(.)$ is a Heaviside step function. This is in agreement with the work of Schiff and Sauer~\cite{schiff2008kalman}. With these assumptions, the model has the form
%\begin{equation}\label{SimplifiedModel}
%\frac{{dv\left( {r,t} \right)}}{{dt}} = \int_\Omega  {w\left( {r-r'} \right)f\left( {v\left( {r',t} \right)} \right)dr'}  - \alpha{v\left( {r,t} \right)} + e\left( {r,t} \right).
%\end{equation}
%In order to make the problem trackable, in terms of model order, we assume that we can decompose the field (membranes voltages) into a set of $n_\phi$ Gaussian basis functions. We will also assume that we know the location and width of the basis function that is required to decompose the field. The state of the system is the height basis functions, rather than the membrane potential. Therefore, the number of basis functions required to decompose the field is equal to the number of states. Also, we assume that we can represent the connectivity by a regular grid of three connectivity Gaussian basis functions (corresponding to short, mid and long range connectivity) of known width. The height of the connectivity basis functions ($\theta$ parameter) is to be estimated.
%\subsection{IDE Formulation}

%For the state-space derivation, to create an estimator, the model can be reformulated like so
%\begin{equation}\label{new_spatiotemporal_kernels}
%\frac{{dv\left( {r,t} \right)}}{{dt}} = \int_\Omega  {k\left( {r - r'} \right)\tilde v\left( {r',t} \right) - \alpha \delta \left( {r - r'} \right)\bar v\left( {r,t} \right)dr'}  + e\left( {r,t} \right)
%\end{equation}
%\begin{equation}\label{define_v_tilde}
%\tilde v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
%   {v\left( {r,t} \right)} & {v\left( {r,t} \right) \ge {v_0}}  \\
%   0 & {otherwise}  \\
%\end{array}} \right.
%\end{equation}
%\begin{equation}\label{define_v_bar}
%\bar v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
%   {v\left( {r,t} \right)} & {v\left( {r,t} \right) < {v_0}}  \\
%   0 & {otherwise}  \\
%\end{array}} \right.
%\end%equation}
% \begin{figure*}
% \centering
% \includegraphics[scale=0.3]{ExampleFieldDecomposition.eps}
% \caption[ExampleFieldDecomposition]{Illustration of a one dimensional field decomposition using 5 Gaussian basis functions, each weighted an element of the state vector. Each basis function has a width of 0.01 and is uniformly positioned with a separation of 0.25 (with arbitrary units).}
% \label{ExampleFieldDecomposition}
% \end{figure*}

%\section{State-Space Neural Field Model}
%In this section we shall reconstruct the model in a state-space form that is a novel representation of the neural field equations. The benefit of restating the model in this format is that it becomes easier to simulate and allows the connectivity to be inferred from data. \begin{equation}
% v_{t+1}(\mathbf r)=\int_{\Omega} \alpha \delta(\mathbf r-\mathbf r')\bar v_t(\mathbf r)d \mathbf r'+\int_{\Omega} w(\mathbf r -\mathbf r')\tilde v_t(\mathbf r')d\mathbf r'+e_t(\mathbf r)
%\label{eq:cortexfield}
%decomposing the field and the kernel we have
%\begin{equation}
%\bar v_t(\mathbf r)=\boldsymbol \phi^{\top}(\mathbf r)\bar{\mathbf x}_t
%\label{eq:cortexfieldlv0}
%\begin{equation}
%\tilde v_t(\mathbf r)=\boldsymbol \phi^{\top}(\mathbf r)\tilde{\mathbf x}_t=\nu_0
%\label{eq:cortexfieldgv0}
%\begin{equation}
% w(\mathbf r -\mathbf r')=\boldsymbol \theta^{\top}\psi(\mathbf r -\mathbf r')
%\label{eq:kerneldecomposition}
%Substituting (\ref{eq:cortexfieldlv0}), (\ref{eq:cortexfieldgv0}) and (\ref{eq:kerneldecomposition}) in (\ref{eq:cortexfield}) we get
%\begin{equation}
% \boldsymbol \phi^{\top}(\mathbf r)\bar{\mathbf x}_{t+1}+\boldsymbol \phi^{\top}(\mathbf r)\tilde{\mathbf x}_{t+1}=
%\int_\Omega\alpha\delta(\mathbf r- \mathbf r')\boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_td\mathbf r'+\int_\Omega\boldsymbol \theta^{\top}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')\tilde{\mathbf x}_td\mathbf r'+e_t(\mathbf r)
%\label{eq:decomposedcortexfield}
%\begin{equation}
% \boldsymbol \phi^{\top}(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\int_{\Omega}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')\tilde{\mathbf x}_td\mathbf r'+e_t(\mathbf r)
%by defining
%\begin{equation}
% \boldsymbol\Phi(\mathbf r)=\int_{\Omega}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')d\mathbf r'
%\label{eq:convolution}
%we have
%\begin{equation}
% \boldsymbol \phi^{\top}(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)\tilde{\mathbf x}_t+e_t(\mathbf r)
%\label{eq:cortexfieldwithPhi}
%Premultiplying (\ref{eq:cortexfieldwithPhi}) by $\boldsymbol\phi(\mathbf r) $, and integrating over $\mathbf r$
%\begin{equation}
%\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \phi^{\top}(\mathbf r)d\mathbf s(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \phi^{\top}(\mathbf r)d\mathbf r \bar{\mathbf x}_{t}+\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)d\mathbf r\tilde{\mathbf x}_t+\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r
%\label{eq:cortexfieldversion2}
%\begin{equation}
% \boldsymbol\Psi_x(\mathbf r)=\int_{\Omega}\boldsymbol\phi(\mathbf r)\boldsymbol\phi^{\top}(\mathbf r)d\mathbf r
%\label{eq:psix}
%\begin{equation}
%  \boldsymbol\Psi_{\theta}(\mathbf r)=\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)d\mathbf r
%\label{eq:psitheta}
%substituting (\ref{eq:psix}) and (\ref{eq:psitheta}) in (\ref{eq:cortexfieldversion2}) we get
%\begin{equation}
%\boldsymbol\Psi_x(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\boldsymbol\Psi_x(\mathbf r) \bar{\mathbf x}_{t}+\boldsymbol\Psi_{\theta}(\mathbf r)\tilde{\mathbf x}_t+\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r
%pre-multiplying both sides by $\boldsymbol\Psi_x^{-1}$
%\begin{equation}
%\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1}=\alpha\bar{\mathbf x}_{t}+\boldsymbol\Psi_x^{-1}\boldsymbol\Psi_{\theta}\tilde{\mathbf x}_t+\boldsymbol\Psi_x^{-1}\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r
%????????????????????????????????
%\subsection{Observation equation}
% \mathbf y_t(\mathbf r)&=&\\
%&=&\int_{\Omega}H(\mathbf r-\mathbf r')v_t(\mathbf r')d\mathbf r'+\boldsymbol\epsilon_t \\
%&=&\int_{\Omega}H(\mathbf r-\mathbf r')\bar v_t(\mathbf r')d\mathbf r'+\int_{\Omega}H(\mathbf r-\mathbf r')\tilde v_t(\mathbf r')d\mathbf r'+\epsilon_t
%\label{eq:observationequation}
% we decompose $ H(\mathbf r-\mathbf r')$ as
% \begin{equation}
%  H(\mathbf r-\mathbf r')= \mathbf c^{\top}\mu(\mathbf r-\mathbf r')
%\label{eq:decomposedobservationkernel}
%and substituting  (\ref{eq:cortexfieldlv0}), (\ref{eq:cortexfieldgv0}) and (\ref{eq:decomposedobservationkernel}) in (\ref{eq:observationequation}) we get
%\begin{equation}
% \mathbf y_t(\mathbf r)=\int_{\Omega}\mathbf c^{\top}\mu(\mathbf r-\mathbf r')\phi^{\top}(\mathbf r')\bar{\mathbf x}_{t}d\mathbf r'+\int_{\Omega}\mathbf c^{\top}\mu(\mathbf r-\mathbf r')\phi^{\top}(\mathbf r')\tilde{\mathbf x}_{t}d\mathbf r'+\epsilon_t
%and finally we have
%\begin{equation}
%\mathbf y_t(\mathbf r)= \begin{bmatrix}\mathbf c^{\top}\mathbf M(\mathbf r)&\mathbf c^{\top}\mathbf M(\mathbf r)\end{bmatrix}\underline{\mathbf x}_t+\epsilon_t
%where
%\begin{equation}
% \mathbf M(\mathbf r)=\int_{\Omega}\mu(\mathbf r-\mathbf r')\phi^{\top}(\mathbf r')d\mathbf r'
%Next we shall demonstrate the stability of the model and in the following section the estimator for the state vector and connectivity structure is derived.
